Determining the alternatives for scalar implicature

### 2016-01-31 Paper/Paper.Rmd
first eyoon edits, need to fix stuhmuller umlaute

How much of what we mean comes from the words that go unsaid? As listeners, our ability to make precise inferences about a speaker's intended meaning in context is  indispensable to successful communication. For example, listeners commonly enrich the meaning of the scalar item *some* to *some but not all* in sentences like "Alice ate some of the cookies" [@grice1975;@horn1984;@levinson2000]. These inferences, called _scalar implicatures_, have been an important test case for understanding pragmatic inferences more generally. A Gricean account of this phenomenon assumes listeners reason about the meaning the speaker intended by incorporating knowledge about a) alternative scalar items a speaker could have used (such as *all*) and b) the relative informativity of using such alternatives [@grice1975]. According to this account, a listener will infer that the speaker must have intended that Alice did not eat all the cookies because otherwise it would have been underinformative to use the descriptor *some* when the alternative *all* could have been used just as easily.

@@

@franke2014 used an RSA-style model to assess what alternatives a speaker would need to consider in order to produce the typicality/felicity ratings reported by @degen2015 and @vantiel2014. In order to do this, @franke2014's model assigned weights to a set of alternative numerical expressions. Surprisingly, along with weighting *one* highly (a conclusion that was supported by the empirical work), the best-fitting model assigned substantial weight to *none* as an alternative. This finding was especially surprising considering the emphasis of standard theories on scalar items that stand in entailment relationships with one another (e.g. *one* entails *some* even if it is not classically considered to be part of the scale).

>>>>

How much of what we mean comes from the words that go unsaid? As listeners, our ability to make precise inferences about a speaker's intended meaning in context is  indispensable to successful communication. For example, listeners commonly enrich the meaning of the scalar item *some* to *some but not all* in sentences like "Alice ate some of the cookies" [@grice1975;@horn1984;@levinson2000]. These inferences, called _scalar implicatures_, have been an important test case for understanding pragmatic inferences more generally. A Gricean account of this phenomenon assumes listeners reason about a speaker's intended meaning by incorporating knowledge about a) alternative scalar items a speaker could have used (such as *all*) and b) the relative informativity of using such alternatives [@grice1975]. According to this account, a listener will infer that the speaker must have intended that Alice did not eat all the cookies because it would have been underinformative to use the descriptor *some* when the alternative *all* could have been used just as easily.

@@

@franke2014 used an RSA-style model to assess what alternatives a speaker would need to consider in order to produce the typicality/felicity ratings reported by @degen2015 and @vantiel2014 for the scale *some*/*all*. In order to do this, @franke2014's model assigned weights to a set of alternative numerical expressions. Surprisingly, along with weighting *one* highly (a conclusion that was supported by the empirical work), the best-fitting model assigned substantial weight to *none* as an alternative to *some*. This finding was especially surprising considering the emphasis of standard theories on scalar items that stand in entailment relationships with one another (e.g. *one* entails *some* even if it is not classically considered to be part of the scale).

### 2016-01-31 Paper/Paper.tex
first eyoon edits, need to fix stuhmuller umlaute

How much of what we mean comes from the words that go unsaid? As
listeners, our ability to make precise inferences about a speaker's
intended meaning in context is indispensable to successful
communication. For example, listeners commonly enrich the meaning of the
scalar item \emph{some} to \emph{some but not all} in sentences like
``Alice ate some of the cookies'' (Grice, 1975; Horn, 1984; Levinson,
2000). These inferences, called \emph{scalar implicatures}, have been an
important test case for understanding pragmatic inferences more
generally. A Gricean account of this phenomenon assumes listeners reason
about the meaning the speaker intended by incorporating knowledge about
a) alternative scalar items a speaker could have used (such as
\emph{all}) and b) the relative informativity of using such alternatives
(Grice, 1975). According to this account, a listener will infer that the
speaker must have intended that Alice did not eat all the cookies
because otherwise it would have been underinformative to use the
descriptor \emph{some} when the alternative \emph{all} could have been
used just as easily.

@@

By formalizing pragmatic reasoning, computational models can help
provide more direct evidence about the role that alternatives play. The
``rational speech act'' model (RSA) is one recent framework for
understanding inferences about meaning in context (Frank \& Goodman,
2012; N. D. Goodman \& Stuhlm{ü}ller, 2013). RSA models frame language
understanding as a special case of social cognition, in which listeners
and speakers reason recursively about one another's goals. In the case
of scalar implicature, a listener makes a probabilistic inference about
what the speaker's most likely communicative goal was, given that she
picked the quantifier \emph{some} rather than the stronger quantifier
\emph{all}. In turn, the speaker reasons about what message would best
convey her intended meaning to the listener, given that he is reasoning
in this way. This recursion is grounded in a ``literal'' listener who
reasons only according to the basic truth-functional semantics of the
language.

Franke (2014) used an RSA-style model to assess what alternatives a
speaker would need to consider in order to produce the
typicality/felicity ratings reported by Degen \& Tanenhaus (2015) and
Tiel (2014). In order to do this, Franke (2014)'s model assigned weights
to a set of alternative numerical expressions. Surprisingly, along with
weighting \emph{one} highly (a conclusion that was supported by the
empirical work), the best-fitting model assigned substantial weight to
\emph{none} as an alternative. This finding was especially surprising
considering the emphasis of standard theories on scalar items that stand
in entailment relationships with one another (e.g. \emph{one} entails
\emph{some} even if it is not classically considered to be part of the
scale).

@@

Goodman, N. D., \& Stuhlm{ü}ller, A. (2013). Knowledge and implicature:
Modeling language understanding as social cognition. \emph{Topics in
Cognitive Science}, \emph{5}(1), 173--184.

>>>>

How much of what we mean comes from the words that go unsaid? As
listeners, our ability to make precise inferences about a speaker's
intended meaning in context is indispensable to successful
communication. For example, listeners commonly enrich the meaning of the
scalar item \emph{some} to \emph{some but not all} in sentences like
``Alice ate some of the cookies'' (Grice, 1975; Horn, 1984; Levinson,
2000). These inferences, called \emph{scalar implicatures}, have been an
important test case for understanding pragmatic inferences more
generally. A Gricean account of this phenomenon assumes listeners reason
about a speaker's intended meaning by incorporating knowledge about a)
alternative scalar items a speaker could have used (such as \emph{all})
and b) the relative informativity of using such alternatives (Grice,
1975). According to this account, a listener will infer that the speaker
must have intended that Alice did not eat all the cookies because it
would have been underinformative to use the descriptor \emph{some} when
the alternative \emph{all} could have been used just as easily.

@@

By formalizing pragmatic reasoning, computational models can help
provide more direct evidence about the role that alternatives play. The
``rational speech act'' model (RSA) is one recent framework for
understanding inferences about meaning in context (Frank \& Goodman,
2012; N. D. Goodman \& Stuhlmuller, 2013). RSA models frame language
understanding as a special case of social cognition, in which listeners
and speakers reason recursively about one another's goals. In the case
of scalar implicature, a listener makes a probabilistic inference about
what the speaker's most likely communicative goal was, given that she
picked the quantifier \emph{some} rather than the stronger quantifier
\emph{all}. In turn, the speaker reasons about what message would best
convey her intended meaning to the listener, given that he is reasoning
in this way. This recursion is grounded in a ``literal'' listener who
reasons only according to the basic truth-functional semantics of the
language.

Franke (2014) used an RSA-style model to assess what alternatives a
speaker would need to consider in order to produce the
typicality/felicity ratings reported by Degen \& Tanenhaus (2015) and
Tiel (2014) for the scale \emph{some}/\emph{all}. In order to do this,
Franke (2014)'s model assigned weights to a set of alternative numerical
expressions. Surprisingly, along with weighting \emph{one} highly (a
conclusion that was supported by the empirical work), the best-fitting
model assigned substantial weight to \emph{none} as an alternative to
\emph{some}. This finding was especially surprising considering the
emphasis of standard theories on scalar items that stand in entailment
relationships with one another (e.g. \emph{one} entails \emph{some} even
if it is not classically considered to be part of the scale).

@@

Goodman, N. D., \& Stuhlmuller, A. (2013). Knowledge and implicature:
Modeling language understanding as social cognition. \emph{Topics in
Cognitive Science}, \emph{5}(1), 173--184.
